Quantum Mechanics Assignment One

### 2016-01-20 classical/latex/a9.tex

\textbf{A four particles system consists of masses $m_i$ at coordinates:}
\begin{align*}
  m_1 &= 5m\\
  \vec{r_1} &= (a, 2a, 0)\\
  m_2 &= 5m\\
  \vec{r_2} &= (0, -2a, a)\\
  m_3 &= 5m\\
  \vec{r_3} &= (a, a, -a)\\
  m_4 &= 5m\\
  \vec{r_4} &= (3a, a, -2a)\\
\end{align*}
\textbf{Find the inertia tensor, the principal axes, and the moments of inertia. Explicitly check that the principal axes are orthogonal to each other. (Hint: you can use an online solver for the third degree polynomial, such as: http://www.calculatorsoup.com/calculators/algebra/cubicequation.php)}

>>>>

\textbf{A four particles system consists of masses $m_i$ at coordinates:}
\begin{align*}
  m_1 &= 5m\\
  \vec{r_1} &= (a, 2a, 0)\\
  m_2 &= 2m\\
  \vec{r_2} &= (0, -2a, a)\\
  m_3 &= 3m\\
  \vec{r_3} &= (a, a, -a)\\
  m_4 &= m\\
  \vec{r_4} &= (3a, a, -2a)\\
\end{align*}
\textbf{Find the inertia tensor, the principal axes, and the moments of inertia. Explicitly check that the principal axes are orthogonal to each other. (Hint: you can use an online solver for the third degree polynomial, such as: http://www.calculatorsoup.com/calculators/algebra/cubicequation.php)}

\begin{align*}
  I &=
  \begin{pmatrix}
    \int_V \rho(y^2+z^2)dV & -\int_V xydV & -\int_V xzdV\\
     -\int_V xydV & \int_V \rho(x^2+z^2)dV & -\int_V yzdV\\
     -\int_V xzdV & -\int_V yzdV & \int_V \rho(x^2+y^2)dV\\
  \end{pmatrix}\\
  I &=
  \begin{pmatrix}
    M\sum (y_i^2+z_i^2) & -\sum x_iy_i & -\sum x_iz_i\\
     -\sum x_iy_i & M\sum (x_i^2+z_i^2) & -\sum y_iz_i\\
     -\sum x_iz_i & -\sum y_iz_i & M\sum (x_i^2+y_i^2)\\
  \end{pmatrix}\\
  I &=
  Ma^2
  \begin{pmatrix}
    41 & -16 & 9\\
    -16 & 26 & 9\\
    9 & 9 & 49\\
  \end{pmatrix}
\end{align*}

We now want to find the solutions to the eigenvalue equation:

\begin{align*}
  &\det
  \begin{pmatrix}
    Ma^241-\lambda & -Ma^216 & Ma^29\\
    -Ma^216 & Ma^226-\lambda & Ma^29\\
    Ma^29 & Ma^29 & Ma^249-\lambda\\
  \end{pmatrix} = 0\\
  \lambda_1 &= 54.9631\hspace{0.1cm}Ma^2\\
  \lambda_2 &= 49.3639\hspace{0.1cm}Ma^2\\
  \lambda_3 &= 11.673\hspace{0.1cm}Ma^2\\
\end{align*}

These are our moments of inertia. We can then find our angular eigenvectors, which represent our principal axes:

\begin{align*}
  \omega_1 &=
  \begin{pmatrix}
    0.62\\
    -0.1\\
    0.78\\
  \end{pmatrix}\hspace{1cm}
  \omega_2 =
  \begin{pmatrix}
    -0.59\\
    0.6\\
    0.54\\
  \end{pmatrix}\hspace{1cm}
  \omega_3 =
  \begin{pmatrix}
    -0.53\\
    -0.79\\
    0.32\\
  \end{pmatrix}
\end{align*}

We now verify this is an orthogonal basis:

\begin{align*}
  \omega_1 \cdot \omega_2 = \omega_1 \cdot \omega_3 = \omega_2 \cdot \omega_3 \approx 0\\
\end{align*}

This was a cool problem. I think it's really interesting that all objects have three principal axes. This is not at all intuitive. I can see that all objects have at least one principal axis. And by this, there must be two vectors orthogonal to it, so it makes sense given that assumption that there are three principal axes. This is still unintuitive, though. This problem involved using linear algebra solvers online since the actual eigenvalues were so ugly. But it was a good exercise. It was interesting to take the continuous form of the I equation and discretize it.\\

### 2016-01-20 classical/latex/cm-equations.tex

\section{Problem One: Inertia tensor of 4-particle system}

\textbf{A four particles system consists of masses $m_i$ at coordinates:}
\begin{align*}
  m_1 &= 5m\\
  \vec{r_1} &= (a, 2a, 0)\\
  m_2 &= 2m\\
  \vec{r_2} &= (0, -2a, a)\\
  m_3 &= 3m\\
  \vec{r_3} &= (a, a, -a)\\
  m_4 &= m\\
  \vec{r_4} &= (3a, a, -2a)\\
\end{align*}
\textbf{Find the inertia tensor, the principal axes, and the moments of inertia. Explicitly check that the principal axes are orthogonal to each other. (Hint: you can use an online solver for the third degree polynomial, such as: http://www.calculatorsoup.com/calculators/algebra/cubicequation.php)}

\begin{align*}
  I &=
  \begin{pmatrix}
    \int_V \rho(y^2+z^2)dV & -\int_V xydV & -\int_V xzdV\\
     -\int_V xydV & \int_V \rho(x^2+z^2)dV & -\int_V yzdV\\
     -\int_V xzdV & -\int_V yzdV & \int_V \rho(x^2+y^2)dV\\
  \end{pmatrix}\\
  I &=
  \begin{pmatrix}
    M\sum (y_i^2+z_i^2) & -\sum x_iy_i & -\sum x_iz_i\\
     -\sum x_iy_i & M\sum (x_i^2+z_i^2) & -\sum y_iz_i\\
     -\sum x_iz_i & -\sum y_iz_i & M\sum (x_i^2+y_i^2)\\
  \end{pmatrix}\\
  I &=
  Ma^2
  \begin{pmatrix}
    41 & -16 & 9\\
    -16 & 26 & 9\\
    9 & 9 & 49\\
  \end{pmatrix}
\end{align*}

We now want to find the solutions to the eigenvalue equation:

\begin{align*}
  &\det
  \begin{pmatrix}
    Ma^241-\lambda & -Ma^216 & Ma^29\\
    -Ma^216 & Ma^226-\lambda & Ma^29\\
    Ma^29 & Ma^29 & Ma^249-\lambda\\
  \end{pmatrix} = 0\\
  \lambda_1 &= 54.9631\hspace{0.1cm}Ma^2\\
  \lambda_2 &= 49.3639\hspace{0.1cm}Ma^2\\
  \lambda_3 &= 11.673\hspace{0.1cm}Ma^2\\
\end{align*}

These are our moments of inertia. We can then find our angular eigenvectors, which represent our principal axes:

\begin{align*}
  \omega_1 &=
  \begin{pmatrix}
    0.62\\
    -0.1\\
    0.78\\
  \end{pmatrix}\hspace{1cm}
  \omega_2 =
  \begin{pmatrix}
    -0.59\\
    0.6\\
    0.54\\
  \end{pmatrix}\hspace{1cm}
  \omega_3 =
  \begin{pmatrix}
    -0.53\\
    -0.79\\
    0.32\\
  \end{pmatrix}
\end{align*}

We now verify this is an orthogonal basis:

\begin{align*}
  \omega_1 \cdot \omega_2 = \omega_1 \cdot \omega_3 = \omega_2 \cdot \omega_3 \approx 0\\
\end{align*}

This was a cool problem. I think it's really interesting that all objects have three principal axes. This is not at all intuitive. I can see that all objects have at least one principal axis. And by this, there must be two vectors orthogonal to it, so it makes sense given that assumption that there are three principal axes. This is still unintuitive, though. This problem involved using linear algebra solvers online since the actual eigenvalues were so ugly. But it was a good exercise. It was interesting to take the continuous form of the I equation and discretize it.\\

>>>>

\begin{align*}
  \tau \frac{\partial\phi}{\partial t} &= -\frac{\delta G}{\delta \phi}\\
  \\
  \frac{\partial \tilde{u}}{\partial t} &= \nabla^2 u + F\frac{\partial \phi}{\partial t}\\
  \\
  G(\phi_{i,j}) &= V(\phi_{i,j}) + \frac{\delta^2(\nabla\phi_{i,j})^2}{2}\\
\end{align*}

### 2016-01-20 quantum/latex/a1.tex
Add everything.
add a1.work
add all

\section{Problem 5.11}
This problem deals with an infinite square well wave which suddenly has its right potential wall
move from $L$ to $3L$. We are asked to find the probability of finding the wave in the ground or
first excited state. This problem is basically asking us to find the $c_1$ and $c_2$ coefficients
of the old wave function in the new well, and from them find probabilities.

@@

\section{8.7}

>>>>

\section{Problem 5.11}
This problem asks us to find the probabilities to find $n=2$ Hydrogen electrons in the classically
forbidden region.\\

This problem deals with an infinite square well wave which suddenly has its right potential wall
move from $L$ to $3L$. We are asked to find the probability of finding the wave in the ground or
first excited state. This problem is basically asking us to find the $c_1$ and $c_2$ coefficients
of the old wave function in the new well, and from them find probabilities.

@@

\section{8.7}
We begin with the potential and find the r-values at which $E=V$. We then calculate
the probability using $P = \int_r^\infty \psi^* \psi dr$.

\begin{align*}
  V(r) &= \frac{-e^2}{4\pi\epsilon_0r}\\
\end{align*}

We first solve for the forbidden distance:

\begin{align*}
  V(r) &= E_2\\
  -\frac{Z\alpha\hslash c}{r} &= -\frac{1}{2n^2}\alpha^2m_ec^2\\
  \frac{-1.44 \mbox{eV nm}}{r} &= -3.4 \mbox{eV}\\
  r &= 2.36 \mbox{nm}
\end{align*}

Next we find the probability of our particles being further than this:

\begin{align*}
  P_{200} &= \int_{2.36}^\infty \int_0^{2\pi} \int_0^\pi
  \frac{r^2}{32\pi a_0^3}\left(2-\frac{r}{a_0}\right)^2e^{\frac{-r}{a_0}}
  \sin(\theta)drd\phi d\theta = 2.09*10^{-14}\\
  P_{21-1} &= \int_{2.36}^\infty \int_0^{2\pi} \int_0^\pi
  \frac{r^4}{64\pi a_0^5}e^{\frac{-r}{a_0}}\sin^3(\theta)
  drd\phi d\theta = 7.62*10^{-15}\\
  P_{210} &= \int_{2.36}^\infty \int_0^{2\pi} \int_0^\pi
  \frac{r^4}{32\pi a_0^5}e^{\frac{-r}{a_0}}\cos^2(\theta)
  \sin(\theta)drd\phi d\theta = 7.62*10^{-15}\\
  P_{211} &= \int_{2.36}^\infty \int_0^{2\pi} \int_0^\pi
  \frac{r^4}{64\pi a_0^5}e^{\frac{-r}{a_0}}\sin^3(\theta)
  drd\phi d\theta = 7.62*10^{-15}\\
\end{align*}

We note that it is much more likely for the s-shell electron to be found in the forbidden (far
from atom) zone than the p-shell electrons. This is indicative of the shape of the s vs p shells
and how spread-out they are from the nucleus. We also note that, though we started off with 3
different wave functions for the $\ell = 1$ states, we get the same probabilities in the end.
This is to be expected, since the different $m_\ell$-numbered states are just rotations of each
other around the atom's axis, their shapes are the same.\\

I was initially confused as to why the units of the wave function were in $m^{-3/2}$, but then
recalled that complex conjugating the wave function and integrating over three dimensions would
cancel out these units.\\
